### 2018-01-12 submission/manuscript.Rmd
update discussion paragraph on b frag and strep gallo

The tissue studies did not provide a clearer understanding of how the microbiota may be associated with lesion. Generally, the full OTU-based models of unmatched and matched [Figure S7E & F] tissue samples were concordant with stool showing that resident microbes were the most prevalent in the top 10 across study. Unlike in stool, *Fusobacterium* was the only mouth-associated bacteria consistently present in the top 10 of the full CRC models [Figure S7B-C & E-F]. The majority of the tissue results seem to be study specific with many top 10 taxa being present only in a single study. Additionally, the presence of genera associated with contamination, within the top 10 most important variables for the genera and OTU models, is worrying because it is commonly a marker of contamination. The low bacterial biomass of tissue samples coupled with potential contaminationcould ultimately explain why the these results seem to be more sporadic than the stool results.

One important caveat to this study is that even though specific genera associated with species such as *Bacteroides fragilis* and *Streptococcus gallolyticus* subsp. *gallolyticus* were not identified, it does not necessarily mean these specific species are not important in human CRC [@bfrag_Boleij_2014; @strep_Aymeric_2017]. Since we are limited in our aggregation of the data, across study, to the genus level, it is not possible to clearly delineate what specific species are contributing to overall disease progression. However, *Streptococcus gallolyticus* subsp. *gallolyticus* is a mouth-associated micobe and our study clearly shows that regardless of whether the sample is tissue or stool, mouth-associated genera are commonly associated with an increased RR of both adenoma and carcinoma.

>>>>

The tissue studies did not provide a clearer understanding of how the microbiota may be associated with lesion. Generally, the full OTU-based models of unmatched and matched [Figure S7E & F] tissue samples were concordant with stool showing that resident microbes were the most prevalent in the top 10 across study. Unlike in stool, *Fusobacterium* was the only mouth-associated bacteria consistently present in the top 10 of the full CRC models [Figure S7B-C & E-F]. The majority of the tissue results seem to be study specific with many top 10 taxa being present only in a single study. Additionally, the presence of genera associated with contamination, within the top 10 most important variables for the genera and OTU models, is worrying because it is commonly a marker of contamination. The low bacterial biomass of tissue samples coupled with potential contamination could ultimately explain why the these results seem to be more sporadic than the stool results.

One important caveat to this study is that even though specific genera associated with species such as *Bacteroides fragilis* and *Streptococcus gallolyticus* subsp. *gallolyticus* were not identified, it does not necessarily mean these specific species are not important in human CRC [@bfrag_Boleij_2014; @strep_Aymeric_2017]. Since we are limited in our aggregation of the data, across study, to the genus level, it is not possible to clearly delineate what specific species are contributing to overall disease progression. Our observations are not inconsistent with the previous literature on either *Bacteroides fragilis* or *Streptococcus gallolyticus* subsp. *gallolyticus*. As an example, for stool, the full RF models consistently identified the genus *Bacteroides* as well as OTUs that classified as *Bacteroides* to be important model component across studies. Suggesting that even though *Bacteroides* may not increase the RR of lesion and may not vary in relative abundance, like *Fusobacterium*, it is still important in CRC. Additionaly, *Streptococcus gallolyticus* subsp. *gallolyticus* is a mouth-associated micobe and our study clearly shows that regardless of whether the sample is tissue or stool, mouth-associated genera are commonly associated with an increased RR of both adenoma and carcinoma.

### 2018-01-12 submission/manuscript.tex
update discussion paragraph on b frag and strep gallo

The tissue studies did not provide a clearer understanding of how the
microbiota may be associated with lesion. Generally, the full OTU-based
models of unmatched and matched {[}Figure S7E \& F{]} tissue samples
were concordant with stool showing that resident microbes were the most
prevalent in the top 10 across study. Unlike in stool,
\emph{Fusobacterium} was the only mouth-associated bacteria consistently
present in the top 10 of the full CRC models {[}Figure S7B-C \& E-F{]}.
The majority of the tissue results seem to be study specific with many
top 10 taxa being present only in a single study. Additionally, the
presence of genera associated with contamination, within the top 10 most
important variables for the genera and OTU models, is worrying because
it is commonly a marker of contamination. The low bacterial biomass of
tissue samples coupled with potential contaminationcould ultimately
explain why the these results seem to be more sporadic than the stool
results.

One important caveat to this study is that even though specific genera
associated with species such as \emph{Bacteroides fragilis} and
\emph{Streptococcus gallolyticus} subsp. \emph{gallolyticus} were not
identified, it does not necessarily mean these specific species are not
important in human CRC {[}20,22{]}. Since we are limited in our
aggregation of the data, across study, to the genus level, it is not
possible to clearly delineate what specific species are contributing to
overall disease progression. However, \emph{Streptococcus gallolyticus}
subsp. \emph{gallolyticus} is a mouth-associated micobe and our study
clearly shows that regardless of whether the sample is tissue or stool,
mouth-associated genera are commonly associated with an increased RR of
both adenoma and carcinoma.

>>>>

The tissue studies did not provide a clearer understanding of how the
microbiota may be associated with lesion. Generally, the full OTU-based
models of unmatched and matched {[}Figure S7E \& F{]} tissue samples
were concordant with stool showing that resident microbes were the most
prevalent in the top 10 across study. Unlike in stool,
\emph{Fusobacterium} was the only mouth-associated bacteria consistently
present in the top 10 of the full CRC models {[}Figure S7B-C \& E-F{]}.
The majority of the tissue results seem to be study specific with many
top 10 taxa being present only in a single study. Additionally, the
presence of genera associated with contamination, within the top 10 most
important variables for the genera and OTU models, is worrying because
it is commonly a marker of contamination. The low bacterial biomass of
tissue samples coupled with potential contamination could ultimately
explain why the these results seem to be more sporadic than the stool
results.

One important caveat to this study is that even though specific genera
associated with species such as \emph{Bacteroides fragilis} and
\emph{Streptococcus gallolyticus} subsp. \emph{gallolyticus} were not
identified, it does not necessarily mean these specific species are not
important in human CRC {[}20,22{]}. Since we are limited in our
aggregation of the data, across study, to the genus level, it is not
possible to clearly delineate what specific species are contributing to
overall disease progression. Our observations are not inconsistent with
the previous literature on either \emph{Bacteroides fragilis} or
\emph{Streptococcus gallolyticus} subsp. \emph{gallolyticus}. As an
example, for stool, the full RF models consistently identified the genus
\emph{Bacteroides} as well as OTUs that classified as \emph{Bacteroides}
to be important model component across studies. Suggesting that even
though \emph{Bacteroides} may not increase the RR of lesion and may not
vary in relative abundance, like \emph{Fusobacterium}, it is still
important in CRC. Additionaly, \emph{Streptococcus gallolyticus} subsp.
\emph{gallolyticus} is a mouth-associated micobe and our study clearly
shows that regardless of whether the sample is tissue or stool,
mouth-associated genera are commonly associated with an increased RR of
both adenoma and carcinoma.
